---
title: "home5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Frequency Analysis

The data SMOKING1.txt contains a 2-dimensional frequency table, where the em-
ployees of a firm have been categorized according to their position (5 categories: SM
= Senior Managers, JM= Junior Managers, SE = Senior Employees, JE = Junior
Employees, SC = Secretaries). Furthermore, the smoking of the employees have 4
categories (None, Light, Medium, Heavy).

## Q1

Form the theoretical frequencies under independence.

```{r}
tb <- read.table("data/SMOKING.txt", header = T, row.names = 1)
#dim(tb)
tb <- as.matrix(tb[-6,-5])
#View(tb)

v1 <- margin.table(tb,1)
v2 <- margin.table(tb,2)

V1 <- matrix(v1,ncol=1) 
V2 <- matrix(v2,nrow=1) 

E <- V1 %*% V2 /sum(tb)
E

```

## Q2

Calculate the attraction repulsion matrix.

```{r}
AR.matrix <- tb/E  #D = original data (number of observations), 
#E = expected number of observations under independence

AR.matrix
```

## Q3

Is smoking more frequent among a specific group in this company?

```{r}
library(ggplot2)
library(reshape)

melted <- melt(AR.matrix)
#View(melted)
#range(melted$value)
melted$X2 <- factor(melted$X2, levels = c("None", "Light","Medium", "Heavy"))
ggplot(melted, aes(x=X1, y=X2, fill=value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = median(melted$value), limit = c(0.5, 1.88),
                       name="AR value")
```
